### 2011-06-04 docs/pubs/EIM-2011/application.tex
resolve section references from intro.tex to other .tex files
use two columns for screen shots. crop images rescale diagram in omnigraffle
scale to 1.0 for omni diagram

\section{Discovery and Integration}

\subsection{Concept query}
The semantic query interface primarily supports locating datasets by how well their observational models match the given criteria. Query criteria largely mirror the structure of an Annotation in that combinations of Entity, Characteristic and Protocol are specified and optionally compounded when increased precision is sought. 
By leveraging the relationships defined and/or inferred from the ontology we are able to increase recall beyond what is possible for simple keyword-based searches (Berkely et al, 20??). Broad queries return direct matches and also n-depth subclass matches. The queries can be quickly honed when using this browsing paradigm which allows rapid exploration of the data repository without the oneous of defining complete observational queries \emph{de novo}. Measurement templates as defined in OBOE compatible ontologies enable a single concept to proxy its constituent parts, namely the valid Characteristics of certain Entities that can be measured with certain Protocols and Standards. This short-hand query generation highlights one of the compelling applications of using OBOE compatible ontologies and can also be leveraged when authoring Annotations.
Using compound semantic query criteria applies a fine-grained filter on the datasets that are returned. Results may include only those datasets that include measurements for a set of specific Characteristics of a particular Entity. Furthermore, a query can specify that those measurements come from precisely the \emph{same instance of that Entity} as belaboured by the authors of the OBOE specification (Schildhauer here somewhere??).

@@

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/metacat-query.png}
\caption{Semantic data query web interface. Data packages containing observations of Kelp Wet Mass less than or equal to 5 [grams] are returned. Additional search options and compund query criteria can be specified within the other tabs. Matches can be saved in the data cart for later exploration.}
\label{fig:metacat-query}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{images/integration.png}
\caption{Data integration query across multiple data packages (A, B).  Annotations (C, D) are used to determine semantically equivalent data attributes contained in the data objects (E, F). Attributes 'plot' and 'site' are considered equivalent measurements of the characteristic 'Location'; attributes 'weight' and 'wt' both map to the same characteristic 'Mass'. The Semantic Mediation API utilizes the Data Manager Library to load and query the source data informed by semantic similarities between the structurally disparate data attributes.}
\label{fig:integration}
\end{figure}

>>>>

\section{Discovery and Integration}
\label{sec:application}

\subsection{Concept query}
The semantic query interface primarily supports locating datasets by how well their observational models match the given criteria. Query criteria largely mirror the structure of an Annotation in that combinations of Entity, Characteristic and Protocol are specified and optionally compounded when increased precision is sought.

By leveraging the relationships defined and/or inferred from the
ontology we are able to increase recall beyond what is possible for
simple keyword-based searches
\cite{berkley09:_improv_data_discov_for_metad}. Broad queries return
direct matches and also n-depth subclass matches. The queries can be
quickly honed when using this browsing paradigm which allows rapid
exploration of the data repository without the oneous of defining
complete observational queries \emph{de novo}. Measurement templates
as defined in OBOE compatible ontologies enable a single concept to
proxy its constituent parts, namely the valid Characteristics of
certain Entities that can be measured with certain Protocols and
Standards. This short-hand query generation highlights one of the
compelling applications of using OBOE compatible ontologies and can
also be leveraged when authoring Annotations.  Using compound semantic
query criteria applies a fine-grained filter on the datasets that are
returned. Results may include only those datasets that include
measurements for a set of specific Characteristics of a particular
Entity. Furthermore, a query can specify that those measurements come
from precisely the \emph{same instance of that Entity} as belaboured
by the authors of the OBOE specification (Schildhauer here
somewhere??).

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=1.0\textwidth]{images/metacat-query.png}
\caption{Semantic data query web interface. Data packages containing observations of Kelp Wet Mass less than or equal to 5 [grams] are returned. Additional search options and compund query criteria can be specified within the other tabs. Matches can be saved in the data cart for later exploration.}
\label{fig:metacat-query}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics[scale=1.0]{images/integration.png}
\caption{Data integration query across multiple data packages (A, B).  Annotations (C, D) are used to determine semantically equivalent data attributes contained in the data objects (E, F). Attributes 'plot' and 'site' are considered equivalent measurements of the characteristic 'Location'; attributes 'weight' and 'wt' both map to the same characteristic 'Mass'. The Semantic Mediation API utilizes the Data Manager Library to load and query the source data informed by semantic similarities between the structurally disparate data attributes.}
\label{fig:integration}
\end{figure}

### 2011-06-04 docs/pubs/EIM-2011/related.tex
resolve section references from intro.tex to other .tex files

\section{Introduction}
\label{sec-introduction}

>>>>

\section{Related Work}
\label{sec:related}

The need for uniform mechanisms to describe observational data has led
to many proposals for observational data models (e.g., \cite{om,
  tarboton07:_cuahs_commun_obser_data_model,netcdf}), and ontologies
(e.g.,
\cite{fox09:_ontol,sweet,sheth08:_seman_sensor_web,mungall110:_integ}).
The work here is complementary to these efforts by providing a
framework and extensions to popular metadata tools (namely, Metacat
\cite{} and Morpho \cite{}) for supporting a generic observational
model through relatively light-weight, formal annotations. Our
approach also supports the use of domain-specific ontologies for
describing data as well as high-level query support for discovering
data sets. Our extensions also support ... integration ... tasks that
are crucial for determining the relevance of data.

Data management systems are increasingly employing various forms of
annotations to help improve search (e.g.,
\cite{DBLP:conf/icde/GeertsKM06,Reeve05SemanticAnnotation,DBLP:conf/vldb/Bhagwat
  CTV04}). For example, MONDRIAN \cite{DBLP:conf/icde/GeertsKM06}
employs an annotation model and query operators to manipulate both
data and annotations.  However, users must be familiar with the
underlying data structures (schemas) to take advantage of these
operators, which is generally not feasible for observational data in
which data sets exhibit a high degree of structural and semantic
heterogeneity.  Some efforts have also been carried out for leveraging
annotations, e.g., for the discovery of domain-specific data
\cite{obsdatasearch09,DBLP:conf/icde/StoyanovichMR10}, however, these
approaches are largely based on keyword queries, and do not consider
structured searches. Our work differs in that we consider a highly
structured and generic model for annotations with the aim of providing
a uniform approach for issuing structured data-discovery searches. Our
work is closely aligned to traditional data integration approaches
(e.g., \cite{HalevyDataIntegration06,kolaitis05}), where a global
mediated schema is used to (physically or logically) merge the
structures of heterogeneous data sources using mapping constraints
among the source and target schemas. As such, the observational model
we employ in our framqework can be viewed as a (general-purpose)
mediation schema for observational data sets.  This schema can be
augmented with logic rules (as target constraints) and uses the
semantic annotations as mapping constraints. However, instead of users
specifying logic constraints directly, we provide a high-level
annotation language that simplifies the specification of mappings and
more naturally aligns with the observation model. In addition, our
work focuses on implementing practical approaches for rewriting and
optimizing queries (that can include aggregation and summarization
operators) over our annotation approach.  In particular, our goal is
to create a feasible, scalable, and deployable system for applying
these approaches for data discovery and exploratory data analysis
within existing scientific data repositories.
